---
title: "Introduction to GEDCOM"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is GEDCOM?

Genealogical Data Communication (GEDCOM) is a specification for storing and exchanging genealogical data. It was created by The Church of Jesus Christ of Latter-day Saints in 1984, and has become the de-facto standard used worldwide. The current version is 7.0 and is the version supported by `gedcomS7`.

## What does a GEDCOM file look like?

A GEDCOM file is a plain text file (with extension .ged), which contains *records*. These records contain information on:

* Families
* Individuals
* Multimedia
* Sources
* Repositories
* Notes
* Submitters

There are also two special records that appear at the beginning and end of a file; a header and trailer.

A sample file looks like this:

```{r}
sample_file <- readLines("https://gedcom.io/testfiles/gedcom70/remarriage1.ged")
sample_file
```

Each record is a nested structure containing one or more lines. Lines in a GEDCOM file can have a number of components:

* **Level**: The level in the hierarchical structure. This appears for every line. New records begin at level 0;
* **Cross-reference identifier**: A string (which looks like @XYZ@) that signals the beginning of a new record (apart from header and trailer);
* **Tag**: A short string given immediately after the level or cross-reference identifier that indicates the type of information being provided on the line. These are controlled values. User-defined tags have been allowed in other GEDCOM programs, but they are discouraged here;
* **Line value**: The value associated with the tag. For example, on line 3, the version of the GEDCOM specification that the file adheres to is 7.0. The line value can also be a **cross-reference pointer**, which links to another record in the file (which looks like @XYZ@). In the above example, the Family record beginning on line `r which(sample_file == "0 @F1@ FAM")` references other Individual records who are members of the family.

There will not be any lines that have all of these components. For example, the first line of records do not contain a line value.

The [GEDCOM 7.0 specification](https://gedcom.io/specifications/FamilySearchGEDCOMv7.html) is substantial, describing the strict rules around the specific set of tags and hierarchical structures allowed for each record type. You shouldn't need to know any of the above to use the package, it is designed to automate the construction of these files.

