---
title: "Implementation of gedcomS7"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why gedcomS7?

One of the main characteristics I wanted for this package was to hide the complexity of the GEDCOM specification, and try to automate genealogical tasks that are time consuming to do manually.

I spent a significant amount of time before writing any code considering my options for how the data would be stored under the hood. In one blog I considered storing genealogical data in a relational table format as it is easier to deal with, but discounted it very quickly as it is not well suited to nested data (and list columns are not easy to deal with).

I toyed with the idea of using an off-the-shelf open source product like [GRAMPS](https://gramps-project.org) but I found it awkward to use and wanted something where I was in complete control, taking full advantage of the strengths of R.

I also considered using data structures more suited to this type of data, such as JSON or graphs (using the `igraph` or `data.tree` package). However, I discovered it would be quite difficult representing some of the structures in the GEDCOM specification to my satisfaction.

My preference was for an Object Orientated approach, which is what many other applications use, but none of the OOP solutions in R quite fit the bill.

My first serious attempt at doing this resulted in `tidyged` and the other packages of the `gedcompendium`. This adopted a dataframe and `tidyverse` approach and whilst it did work it was dependency-heavy and required a lot of processing under the hood.

The release of `S7` presented an ideal opportunity to revisit creating an OOP-based GEDCOM package in R. Initial testing had promising results and so I continued building out the rest of the package, using the updated GEDCOM 7.0 specification.

## The gedcomS7 object

The main GEDCOM object is an S7 representation of a GEDCOM file:

```{r}
library(gedcomS7)

ged <- read_gedcom("https://gedcom.io/testfiles/gedcom70/remarriage1.ged")

str(ged)
```

Properties of the GEDCOM object can be accessed and modified using the `@` operator, e.g.

```{r}
ged@default_language
ged@default_language <- "en"
```

This ease of modification of specific properties of a GEDCOM object wasn't possible with the `tidyged` package.

Properties which don't have values are either empty vectors, empty lists, or NULL (depending on the property). Many properties take values which are particular `gedcomS7` objects (or lists of them, if they take more than one). For ease of use, you are often permitted to provide a simple atomic vector, and `gedcomS7` will convert these to their relevant objects automatically, but only if the object only requires its first property. For example, the `@notes` property can take a simple character vector of notes, and these will be converted to a list of `Note()` objects.

```{r}
ged@notes <- c("This is a note", "This is another note")
ged@notes
```

You can then access properties of these:

```{r}
ged@notes[[2]]@language <- "en"
```

Some of these properties are also read-only (calculated from other properties), such as `@c_next_xref` and `@c_as_ged`. The exploration of all properties of the GEDCOM object is beyond the scope of this article, however the implementation of how properties are stored is important.

## The Push/Pull paradigm

A GEDCOM file could contain many thousands of records containing information on individuals, families, notes, sources, etc. Whilst storing each of these records as `S7` objects within the main GEDCOM object is theoretically possible, in practice it very quickly eats up too much memory rendering the idea a non-starter.

For this reason, records are stored in their raw form from the GEDCOM file as lists of character vectors. For example the `@indi` property is a list of individual records. The lines in the GEDCOM file for the first individual are:

```{r}
ged@indi[[1]]
```

You can also reference records by their xref:

```{r}
ged@indi[["@I1@"]]
```

If you want to edit a record, you must first Pull it from the GEDCOM object. This takes a copy of the record and parses it into an editable `S7` object.

```{r}
john_public <- pull_record(ged, "@I1@")

john_public
```


We can edit a property and Push it back to the GEDCOM object:

```{r}
john_public@notes <- "John once had a dog called Rover"

ged <- push_record(ged, john_public)

ged@indi[[1]]
```

You should never attempt to modify the records in their character vector form directly from the GEDCOM object - additional checks and automated tasks are carried out to ensure self-consistency during the Push process.
