---
title: "Creation of GEDCOM objects"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Creating new objects

The easiest way to create a GEDCOM object is to import an existing GEDCOM file. When you `print()` or use `summary()` on a GEDCOM object it provides a brief summary of the file. The example below imports the smallest legal GEDCOM file:

```{r}
library(gedcomS7)

ged <- read_gedcom("https://gedcom.io/testfiles/gedcom70/minimal70.ged")

summary(ged)
```

A GEDCOM object can also be created using the `new_gedcom()` function:

```{r}
ged_new <- new_gedcom()

ged_new
```

Information about the `gedcomS7` package (as the system/product creating the file) is given in the appropriate places, creation date, as well as a default language (English). The default language can be changed in the `new_gedcom()` function. This must be a character string of language tags as defined in [BCP 47](https://www.rfc-editor.org/info/bcp47) (as must any other language parameter in the package).

Once you have a GEDCOM object in R, you can write it to a valid GEDCOM file using `write_gedcom()`. The filepath must end with file extension .ged.

## GEDCOM object properties

The properties of a GEDCOM object are:

```{r}
str(ged, max.level = 1)
```

The properties of the main GEDCOM object are either:

* Properties of the header in the `@header` property (metadata about the file as a whole).
* Information about records in the `@records` property. This is composed of:
  * Properties that summarise and control the cross-reference identifiers (xrefs) given to records (e.g. `@XREFS`, `@prefixes`, and `@XREFS_NEXT`);
  * Lists of records of each type (`@RAW`);
* Those that allow the user to control whether creation/change dates are added to records as they are pushed to the GEDCOM object (`@update_change_dates` and `@add_creation_dates`) - they don't by default as it can bloat files considerably;
* Those that show you what the GEDCOM file representation looks like (`@GEDCOM`).

Some of these properties are also read-only, which means they are calculated from other properties and cannot be directly set. These are given in capitals.


