---
title: "Records for defining people and relationships"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Information about people and relationships between them are stored in Individual records and Family records. We will focus on Individual records first, and then move on to Family records.

## Individuals

```{r}
library(gedcomS7)

anakin <- IndividualRecord(
  sex = "M"
)
```

The sex of the individual can be specified with the `sex` property, which is a single letter: M(ale), F(emale), (Interse)X, or U(nknown). If no sex is specified, a value of "U" is used.

### Names

A quick and dirty way of adding names is with a character vector (enclosing forward slashes are GEDCOM's way of indicating a surname):

```{r}
anakin@pers_names <- c("Anakin /Skywalker/", "Darth Vader")
```

Individuals can have more than one name. These can be different kinds of name such as birth name or adoptive name, or if there is uncertainty, variants of these based on conflicting evidence sources. There is also the ability to provide translated variants of names. Given this complexity, a personal name is usually defined using a `PersonalName()` object (or a list of them):

```{r}
anakin@pers_names <- list(
  PersonalName(pers_name = "Anakin /Skywalker/",
               name_type = "BIRTH",
               name_pieces = PersonalNamePieces(given = "Anakin",
                                                surname = "Skywalker"),
               notes = "He had this name since the death of Mace Windu."),
  PersonalName(pers_name = "/Darth/ Vader",
               name_type = "AKA",
               type_phrase = "Dark Side name",
               name_translations = PersonalNameTran(pers_name = "Dark Father",
                                                    language = "en"))
)

anakin@pers_names
```

In the example above we have defined two names in a list of `PersonalName()` objects. "Darth" has been enclosed in forward slashes because it is the closest thing to a family name for Sith Lords!

Name types can also be given (with an optional free text phrase) and must take one of the values of `val_name_types()`:

```{r}
val_name_types()
```

All names should preferably be provided by a set of name pieces, specifying the given name, surname, nickname, etc.

You can also define a name translation with `PersonalNameTran()` or a list of them. There are also opportunities to record notes and source citations for each name defined.

Generally, names are arranged in order of importance; the first name is their primary name (usually birth name).

Once added to the Individual record, the property `@PRIMARY_NAME` will provide the primary (first) name as a simple character string, and `@ALL_NAMES` will provide all names as a simple character vector.

```{r}
anakin@PRIMARY_NAME
anakin@ALL_NAMES
```

### Facts

Attributes and events associated with an individual, such as birth and death, are accessed with the `@facts` property. This can be assigned an `IndividualEvent()` or `IndividualAttribute()` object (or a list of them). These objects allow you to define a fact, together with associated date, cause, place/address, and the age of the individual when the fact applied (among other aspects).

The full list of fact types that can be created are given below:

```{r}
val_individual_attribute_types(inc_generic = TRUE)
val_individual_event_types(inc_generic = TRUE)
```

In the example below, we add 4 facts for Anakin:

```{r}
anakin@facts <- list(
  IndividualAttribute(fact_type = "TITL",
                      fact_val = "Lord",
                      notes = "When he was Darth Vader"),
  IndividualAttribute(fact_type = "NMR",
                      fact_val = "1"),
  IndividualEvent(fact_type = "BIRT",
                  place = "Tatooine"),
  IndividualEvent(fact_type = "DEAT",
                  place = "Death Star",
                  cause = "Injuries sustained from the Emperor's lightning")
)
```

If you add a generic event (EVEN) or attribute (FACT), then as well as defining a fact value, you must also add a fact description which provides a further classification:

```{r}
anakin@facts <- append(
  anakin@facts,
  IndividualEvent(fact_type = "EVEN",
                  fact_val = "Lost legs",
                  fact_desc = "Injury",
                  cause = "Lightsaber duel with Obi-Wan Kenobi",
                  place = "Mustafar")
)
```

You can also assert that an individual never experienced an event via the `@non_events` property. This takes a `NonEvent()` object or a list of them.

```{r}
anakin@non_events <- NonEvent("BARM")
```

A non-event with a date period provided asserts that the event never happened in that period. If no date period is provided, it asserts the event never happened at all.

You can see a summary of all facts associated with an individual with the `df_indi_facts()` function.

```{r}
skywalkers <- new_gedcom() |> 
  push_record(anakin)

df_indi_facts(skywalkers, "@I1@")
```

### Family relationships

One of the most important linkages in an Individual record are with Family records. An individual can either be linked to a family as a spouse or as a child.

Below we create a new empty Family record in our GEDCOM object `skywalkers`, and then link Anakin to this family as a spouse.

```{r}
skywalkers <- push_record(skywalkers, FamilyRecord())

anakin <- pull_record(skywalkers, "@I1@")
anakin@fam_links_spou <- FamilyLinkSpouse("@F1@")
```

When we push this record to the GEDCOM object, it will automatically recognise that a new linkage has appeared, and update the Family record to reflect this:

```{r}
skywalkers <- push_record(skywalkers, anakin)

skywalkers@records@RAW@FAM[["@F1@"]]
```

Creating linkages between individuals and families in this way can get quite cumbersome, so a suite of helper functions to define multiple relationships at once are described in the next section.

## Families

### Members

The key properties of a Family record are for defining the members of the family (`@husb_xref`, `@wife_xref`, `@chil_xrefs`). These are usually given as xref links to Individual records. If they do not have an Individual record then a value of "@VOID@" can be used. However, you will need to provide this as a named vector describing the person, e.g.

```{r}
leia_han <- FamilyRecord()

leia_han@chil_xrefs <- c("Ben Solo" = "@VOID@")
```

A named vector can be used even if an xref to an Individual record is provided.

The Family records do not need to describe traditional marriage unions - they can describe non-marital relationships and same-sex relationships. Children also do not have to be biological. The parameter names are a legacy from previous versions of the GEDCOM specification, which have not yet been updated.

### Automation

A suite of helper functions have been created that allow you to add a spouse, parents, children, or siblings with a single function call. These create the records and linkages automatically, and you can fill in the details later. To help you keep track, the functions also allow you to specify the sexes and/or names of those being added.

As an example, we will add a spouse and two children for Anakin:

```{r}
skywalkers <- skywalkers |> 
  add_spouse("@I1@", sex = "F", spou_name = "Padme /Amidala/",
             fam_xref = "@F1@") |> 
  add_children("@F1@", sexes = "MF", chil_names = c("Luke", "Leia"))
```

These functions allow for quite a lot of automation. The `add_spouse()` function automatically creates an Individual record for Padme with her sex and name, and if there was no existing Family record it would have created that as well. The `add_children()` function creates an Individual record for each of the sexes defined. Names can be optionally provided; if no surnames are included in any names (indicated by forward slashes) then they will be taken from the father (or mother if there is no father). The `add_parents()` and `add_siblings()` functions work in a similar way.

### Facts

Like Individual records, Family records can also have attributes and events associated with them, through the `@facts` property. This can be assigned a `FamilyEvent()` or `FamilyAttribute()` object (or a list of them).

The full list of fact types that can be created are given below:

```{r}
val_family_attribute_types(inc_generic = TRUE)
val_family_event_types(inc_generic = TRUE)
```

There are some types of fact; census and residence, that can apply for Individuals and Families. It is highly recommended you do not use these facts for families, only for individuals. It may intuitively be more efficient to define an event for an entire family, but other GEDCOM parsers can have difficulty with this, and defining these facts for each individual separately is recommended. Despite the name, Family records are actually more about the union of two people, rather than the family unit as a whole, and so the remaining events are related to marriages and divorces.

Like Individual records, events that didn't happen to a family can be defined through the `@non_events` property.

## Links to other records

### Individual records

An alternative way of recording the alter ego of Anakin Skywalker would have been to create a separate Individual record for Darth Vader, and link them together. This is achievable via the `@alia_xrefs` property. Whether you include a one-way, or mutual link is up to you.

Individuals and families may also have associations with individuals that aren't necessarily familial. You can record these with the `@associations` property. This property takes an `Association()` object or a list of them.

This object has a void xref pointer by default, as the person may not have an Individual record, and if this is the case (as in the examples below), an `@indi_phrase` must be given containing their name. A `@relation_is` value is mandatory.

```{r}
anakin <- pull_record(skywalkers, "@I1@")

anakin@associations <- list(
  Association(indi_phrase = "Obi-Wan Kenobi",
              relation_is = "OTHER",
              relation_phrase = "Master",
              notes = "While he was Anakin Skywalker"),
  Association(indi_phrase = "Emperor Palpatine",
              relation_is = "OTHER",
              relation_phrase = "Master",
              notes = "While he was Darth Vader")
)
```

The `@relation_is` property must take one of the following values. If it is "OTHER", then a `@relation_phrase` must be given.

```{r}
val_roles()
```

### Submitter records

You can also link to Submitter records in the following ways:

* Those that have contributed information about the individual/family (`@subm_xrefs`); 
* Those that have an interest in the individual's ancestors (`@anci_xrefs`);
* Those that have an interest in the individual's descendants (`@desi_xrefs`).
