---
title: "Records for evidence and supporting information"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Whilst the Individual and Family records describe the people and relationships, the remaining record types describe the evidence and supporting information. Specifically these are:

* Source records where information was taken from;
* Repository records where source records can be found;
* Multimedia records that describe documents like photos or videos;
* Submitter records that contain information on people who have provided information (i.e. human sources);
* Note records for informal notes that can be referenced in multiple places.

## Source records

Source records are one of the most important types of record in a GEDCOM file. They provide evidence for the assertions made in a file, and require careful management.

```{r}
library(gedcomS7)

birth_source <- SourceRecord(
  facts_recorded = FactsRecorded(
    fact_types = "BIRT, MARR, DEAT",
    date_period = DatePeriod("1900", "2000"),
    territory = "England and Wales"
  ),
  originator = "General Register Office",
  full_title = "BMD Register Book 1900-2000",
  source_text = "This might contain a relevant extract from the register",
  repo_citations = list(
    "@R1@",
    RepositoryCitation("@R2@", call_numbers = "ABC123")
  )
)
```

The `@facts_recorded` property can be optionally provided, containing the facts, time periods, and jurisdictions the source covers. In the example above the source covers births, marriages, and deaths in England and Wales over the period of 1900 to 2000. If deaths were only covered until 1980, then you would provide a list of two `FactsRecorded()` objects - one for births and marriages, and the other for deaths.

The `@repo_citations` property contains links to repository records which details where the source book can be found.

There is an art to deciding the level of granularity in which to express sources. There are generally two approaches:

* ‘Source splitters’ where each specific document has its own Source record;
* ‘Source lumpers’ where an entire class of documents has one Source record.

The best approach will depend on the kind of source document, so a hybrid approach is recommended. This is an [entire topic in itself](https://www.fhug.org.uk/kb/kb-article/citing-sources-method-1-and-method-2/) which we won't dwell on here.

## Repository records

A repository is an archive (such as a library) containing source documents. 

```{r}
library_repo <- RepositoryRecord(
  repo_name = "A library",
  address = Address(
    full = "123 Library Road; Turin Village; Manchester; New Hampshire; 123456; United States of America",
    city = "Manchester",
    state = "New Hampshire",
    postal_code = "123456",
    country = "United States of America"
  ),
  phone_numbers = c("645-618-6578", "645-618-6570", "645-618-6577"),
  emails = c("enquiries@alibrary.com", "enquiries2@alibrary.com"),
  faxes  = "0947393",
  web_pages = "www.alibrary.com"
)
```

## Multimedia records

Multimedia records are used to define a single multimedia object such as a document or a photo. The primary property of this record is the `@files` property which takes a `MediaFile()` object (or a list of them if a group of media files are grouped together).

```{r}
my_family_photo <- MediaRecord(
  files = MediaFile(
    location = "C:/my_files/photo.jpg",
    title = "My Family Photo",
    media_type = "image/jpeg",
    medium = "PHOTO"
  )
)
```

The `@media_type` describes the encoding of the file, as defined in [RFC 2045](https://www.rfc-editor.org/info/rfc2045). A registry of file types is maintained [here](https://www.iana.org/assignments/media-types/media-types.xhtml). See the GEDCOM specification for more details.

The `@medium` property must take one of the following values:

```{r}
unname(val_medium_types())
```

## Submitter records

Submitters are people who have contributed information to the GEDCOM file. The main properties of these records are the names and contact details of these people.

```{r}
informant <- SubmitterRecord(
  subm_name = "Joe Bloggs",
  address = Address("2599 Priory Road; Wells; Somerset; England; BA5 1SH")
)
```

## Note records

Note records allow you to make notes in your file and reference them in multiple places.

```{r}
spelling_note <- NoteRecord(
  text = "Lindsay has alternative spellings of 'Lindsey' and 'Linsey'",
  media_type = "text/plain"
)
```

The `@media_type` can also be of type "text/html" if the `@text` property contains HTML markup (e.g. for new line characters).
