---
title: "Date types"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

As you would expect with a genealogical data format, dates form a major part of information contained within it. Within `gedcomS7` there are a number of different date types that can be defined, and there is a helper function for each type to ensure it is formatted correctly.

## Exact dates

Dates that occur on a specific defined date are defined using `DateExact()`:

```{r}
library(gedcomS7)

DateExact(1999, 7, 5)
DateExact(1956, 12, 8)
DateExact(2008, 4, 1)
```

As well as the GEDCOM formatted strings (printed above and also accessed through the `@GEDCOM_STRING` property), this date object can also be expressed as a date type:

```{r}
DateExact(2008, 4, 1)@as_date
```

## Calendar dates

Calendar dates are the most common type of date used. They can not only be used to create specific dates like `DateExact()`, they underpin the other date types described below. They can be created using `DateGregorian()`:

```{r}
DateGregorian(year = 1999, month = 4, day = 5)
DateGregorian(year = 1999, month = 4)
DateGregorian(year = 1999)
```

Years before the Common Era can be defined using negative years:

```{r}
DateGregorian(year = -105)
```

No other calendars apart from the Gregorian calendar are supported by the package.

## Approximate dates

Approximate dates (i.e. those expressing uncertainty) use the qualifiers 'about', 'calculated', or 'estimated' in conjunction with a `DateGregorian()` object:

```{r}
DateGregorian(year = 1999, month = 4, day = 5) |> DateApprox(about = TRUE)
DateGregorian(year = 1999, month = 4, day = 5) |> DateApprox(calc = TRUE)
DateGregorian(year = 1999, month = 4, day = 5) |> DateApprox(est = TRUE)
```


## Date periods

Date periods can take one or two `DateGregorian()` objects:

```{r}
DatePeriod(start_date = DateGregorian(1956, 7, 26),
           end_date = DateGregorian(1956, 9, 15))
DatePeriod(start_date = DateGregorian(1956, 7, 26))
DatePeriod(end_date = DateGregorian(1956, 9, 15))
```

Providing only one date gives a semi-infinite period.


## Date ranges

Date ranges can be defined using `DateRange()`. Unlike `DatePeriod()`, they describe when an event occurred rather than a duration.

```{r}
DateRange(start_date = DateGregorian(1956, 7, 26),
           end_date = DateGregorian(1956, 9, 15))
DateRange(start_date = DateGregorian(1956, 7, 26))
DateRange(end_date = DateGregorian(1956, 9, 15))
```

It's important to note that approximate dates cannot be used in date ranges or date periods. 

## Date values

Date values can be defined with `DateValue()` and are a higher level of date expression than the elementary values given above. A calendar/approximate/range/period date can be used in conjunction with a date phrase (free text) and time. Instead of being single values, they are GEDCOM substructures:

```{r}
my_date <- DateValue(
  date = DateGregorian(year = 1999, month = 4, day = 5),
  date_phrase = "Someone's birthday",
  time = Time(hour = 13, minute = 28, second = 57)
)

my_date

my_date@GEDCOM
```

It's also possible to only define a DateValue through a phrase:

```{r}
DateValue(date = "", date_phrase = "March 14th")@GEDCOM
```

A `DateSorting()` object is similar to `DateValue()` but it can only take Calendar dates as it is used for sorting.
