---
title: "Implementation of gedcomS7"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Why gedcomS7?

One of the main characteristics I wanted for this package was to hide the complexity of the GEDCOM specification, and try to automate genealogical tasks that are time consuming to do manually.

I spent a significant amount of time before writing any code considering my options for how the data would be stored under the hood. In one blog I considered storing genealogical data in a relational table format as it is easier to deal with, but discounted it very quickly as it is not well suited to nested data (list columns are not ideal for this application).

I toyed with the idea of using an off-the-shelf open source product like [GRAMPS](https://gramps-project.org) but I found it awkward to use and wanted something where I was in complete control, taking full advantage of the strengths of R.

I also considered using data structures more suited to this type of data, such as JSON or graphs (using the `igraph` or `data.tree` package). However, I discovered it would be quite difficult representing some of the structures in the GEDCOM specification to my satisfaction.

My preference was for an Object Orientated approach, but none of the OOP solutions in R quite fit the bill.

My first serious attempt at doing this resulted in `tidyged` and the other packages of the `gedcompendium`. This adopted a `tidyverse` approach and whilst it did work it was dependency-heavy and required a lot of processing under the hood.

The release of `S7` presented an ideal opportunity to revisit creating an OOP-based GEDCOM package in R. Initial testing had promising results and so I continued building out the rest of the package, using the updated GEDCOM 7.0 specification.

## The gedcomS7 object

The `tidyged` object is a tibble representation of a GEDCOM file. Before we see how this is structured, let's see an example of a GEDCOM file:

```{r}
readLines(system.file("extdata", "555SAMPLE.GED", package = "tidyged.io"))
```

Lines in a GEDCOM file can have a number of components:

* **Level**: The level in the hierarchical structure. This appears for every line;
* **Cross-reference identifier**: A string (which looks like @XYZ@) that signals the beginning of a new record (apart from header and trailer);
* **Tag**: A short string given immediately after the level or cross-reference identifier that indicates the type of information being provided on the line. These are controlled values. User-defined tags have been allowed in other GEDCOM programs, but they are discouraged here;
* **Cross-reference pointer**: This links to another record in the file (which looks like @XYZ@). In the above example, the Family Group record beginning on line `r which(readLines(system.file("extdata", "555SAMPLE.GED", package = "tidyged.io")) == "0 @F1@ FAM")` references other Individual records who are members of the family;
* **Line value**: The value associated with the tag. For example, on line 6, the CHARacter encoding for the file is given as UTF-8.

There will not be any lines that have all of these components. For example, the first line of records do not contain a line value. Also, a line will never have a cross-reference pointer and a line value defined. For this reason, the `tidyged` object treats cross-reference pointers as just another line value.

The `tidyged` representation of the above GEDCOM file is given below, using a sample dataset built into the package:

```{r}
knitr::kable(tidyged::sample555)
```

The cross-reference identifier is given in the record column. Event though these aren't given for the header and trailer records they are assigned as "HD" and "TR" respectively in the `tidyged` object.

GEDCOM files can be imported and exported using the `tidyged.io` package. Many GEDCOM processors modify the data on import, perhaps ignoring custom tags. This package reads the file as-is, and the only modification of the file is in line with that recommended in the GEDCOM 5.5.5 specification, specifically; replacing double '@' symbols with single ones, merging CONC/CONT lines, and ensuring appropriate capitalisation of certain values. In general, the package sticks rigidly to the GEDCOM specification and does not use custom tags. There is a real issue with other desktop applications creating their own flavours of file and diverging from the specification; `tidyged` does not and will never do this.

If you want to modify single values in a GEDCOM or do a simple Find/Replace, then this is straightforward to achieve with a simple text editor.

<p style="text-align: right;"><a href="https://jl5000.github.io/tidyged/articles/gedcompendium.html">Next article: The gedcompendium ></a></p>







