---
title: "Identifying relations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Identifying relations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

We identify relations to an individual using the family of `get_*()` functions.

We illustrate their functionality using the family below consisting of three generations: A pair of parents who has a single child (named 'Main Person'). This person forms two separate family groups with two spouses, each resulting in two children.

```{r}
library(tidyged)

three_gen <- gedcom(subm("Me")) |> 
  add_indi(qn = "Parent 1") |> 
  add_indi(qn = "Parent 2") |> 
  add_indi(qn = "Main Person") |> 
  add_indi(qn = "Spouse 1") |> 
  add_indi(qn = "Spouse 2") |>
  add_indi(qn = "Child 1") |> 
  add_indi(qn = "Child 2") |> 
  add_indi(qn = "Child 3") |> 
  add_indi(qn = "Child 4")

p1_xref <- find_indi_name(three_gen, "Parent 1")
p2_xref <- find_indi_name(three_gen, "Parent 2")
main_xref <- find_indi_name(three_gen, "Main")
s1_xref <- find_indi_name(three_gen, "Spouse 1")
s2_xref <- find_indi_name(three_gen, "Spouse 2")
c12_xref <- find_indi_name_all(three_gen, "Child 1|Child 2")
c34_xref <- find_indi_name_all(three_gen, "Child 3|Child 4")

three_gen <- three_gen |>
  add_famg(p1_xref, p2_xref, children = main_xref) |> 
  add_famg(main_xref, s1_xref, children = c12_xref) |>
  add_famg(main_xref, s2_xref, children = c34_xref)

df_famg(three_gen) |> 
  knitr::kable()
```

## Immediate relations

The examples below illustrate the functions.

```{r}
get_indi_partners(three_gen, main_xref) |> describe_records(gedcom = three_gen)
get_indi_parents(three_gen, main_xref) |> describe_records(gedcom = three_gen)
get_indi_children(three_gen, main_xref) |> describe_records(gedcom = three_gen)
get_indi_siblings(three_gen, find_indi_name(three_gen, "Child 1")) |> describe_records(gedcom = three_gen)
get_indi_siblings(three_gen, find_indi_name(three_gen, "Child 1"), inc_half_sibs = TRUE, return_name = TRUE)
get_families_as_child(three_gen, main_xref) |> describe_records(gedcom = three_gen)
get_families_as_partner(three_gen, main_xref) |> describe_records(gedcom = three_gen)
get_famg_partners(three_gen, "@F1@") |> describe_records(gedcom = three_gen)
get_famg_children(three_gen, "@F2@", return_name = TRUE)
```

Many functions include a `birth_only` parameter which allows you to focus on purely biological relationships and/or a `return_name` parameter which allows you to return individual's names instead.

## Branches

One of the more sophisticated features of `tidyged` is the ability to manipulate entire branches of your tree. 

### Descendants

We use the `get_descendants()` function below to identify the descendants of Main Person. By default it will exclude the individual, all spouses, and all associated family groups:

```{r}
get_descendants(three_gen, main_xref)
```

We can use the `describe_records()` function to see descriptions:

```{r}
get_descendants(three_gen, main_xref) |>
  describe_records(gedcom = three_gen)
```

Setting `inc_part = TRUE` will include all partners and their descendants, and all descendants' partners:

```{r}
get_descendants(three_gen, main_xref, inc_part = TRUE) |>
  describe_records(gedcom = three_gen)
```

Setting `inc_indi = TRUE` will include the individual:

```{r}
get_descendants(three_gen, main_xref, inc_indi = TRUE) |>
  describe_records(gedcom = three_gen)
```

Setting `inc_famg = TRUE` will include the individual's families where they are a partner, and all descendants' families:

```{r}
get_descendants(three_gen, main_xref, inc_famg = TRUE) |>
  describe_records(gedcom = three_gen)
```

If we want to remove the individual and their partners, all descendants and their partners, and all families, we can remove them with `remove_records()`:

```{r}
get_descendants(three_gen, main_xref, inc_part = TRUE, inc_indi = TRUE, inc_famg = TRUE) |>
  remove_records(gedcom = three_gen) |> 
  df_famg() |> 
  knitr::kable()
```

This combination can result in the removal of a vast amount of data. It will tell the user precisely what it is removing. Be sure the function has done what you expect before accepting the results. It is recommended that you use this function with extreme caution if you think a descendant (or their spouse) may be connected to an individual on another branch of your tree.

This function will, by default, remove any dead references to records that have not been included in the subsetting.

### Ancestors

We can deal with ancestors in a similar way using the `get_ancestors()` function.

## Automating the creation of Individuals and Family Groups

In all the examples you've seen so far the approach has been to build up the tree one record at a time. There are a number of helper functions that allow you to shortcut this laborious exercise. These functions can create multiple records at once, including Family Group records, where you can go back and add more detail. The functions are:

* `add_parents()`
* `add_siblings()`
* `add_spouse()`
* `add_children()`

They all require the xref of an Individual record (or one to be activated), except for `add_children()`, which requires the xref of a Family Group record. These functions do not change the active record. 

Because of this, you cannot use `add_children()` in a single pipeline with the other functions.

The feedback from these functions gives you the necessary xrefs to then add more detail.

To illustrate, we can build up two families starting with a spouse:

```{r}
from_spou <- gedcom(subm("Me")) |>
  add_indi(sex = "M") |>
  add_parents() |>
  add_siblings(sexes = "MMFF") |>
  add_spouse(sex = "F") 
```

The initial individual (@I1@) gets added as a child to a family (@F1@) with two parents (@I2@ and @I3@) and 4 siblings (@I4@ to @I7@). Finally, he is given a spouse (@I8@) in his own family (@F2@).

Now we have the xref of his family, we can add his two daughters:

```{r}
with_chil <- from_spou |>
  add_children(xref = "@F2@", sexes = "FF")
```
Now we have the records, we can use all of these xrefs to add details like names and facts.



